---
output: pdf_document
fontsize: 10pt
graphics: yes
header-includes:
  - \usepackage{courier}
  - \usepackage{enumitem}
  - \usepackage{setspace}
  - \usepackage{amsmath,amssymb,mathtools,courier,bm,upgreek}
  - \newcommand{\benum}{\begin{enumerate}[leftmargin=*]}
  - \newcommand{\benumr}{\begin{enumerate}[resume,leftmargin=*]}
  - \newcommand{\eenum}{\end{enumerate}}
  - \newcommand{\mbf}{\mathbf}
  - \usepackage{array}
  - \usepackage{graphicx}
  - \usepackage{arydshln}
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage[super]{nth}
  - \usepackage{fancyhdr}
  - \usepackage{lastpage}
  - \usepackage{lipsum}
  - \usepackage{listings}
  - \usepackage{titling}
  - \usepackage{xcolor}
  - \definecolor{ForestGreen}{rgb}{0.13, 0.55, 0.13}
  - \definecolor{RoyalBlue}{rgb}{0.25, 0.41, 0.88}
  - \definecolor{YellowGreen}{rgb}{0.6, 0.8, 0.2}
  - \fancyhead[LH]{City of Boulder}
  - \fancyhead[CH]{Exploratory Data Analysis}
  - \fancyhead[RH]{Justin, Arlyn, and Ocean}
  - \fancyfoot[RF]{\thepage\ of \pageref{LastPage}}
  - \fancyfoot[CF]{}
  - \setlength{\headheight}{13pt}
  - \lstset{language=R, basicstyle=\small\ttfamily, numbersep=5pt,backgroundcolor=\color{white},showspaces=false,showstringspaces=false,showtabs=false,mathescape=false, escapeinside={(*@}{@*)},rulecolor=\color{black}, tabsize=2,captionpos=b,breaklines=true,breakatwhitespace=false,keywordstyle=\textcolor{RoyalBlue}, commentstyle=\textcolor{YellowGreen},stringstyle=\textcolor{ForestGreen}} 
  - \usepackage[many,theorems]{tcolorbox}
  - \newtcolorbox{msol}{top=5pt,box align=center,valign=center,colback=white,colframe=blue}
  - \newcommand{\bmsol}{\begin{msol}}
  - \newcommand{\emsol}{\end{msol}}
---

```{r setup, include=FALSE}
#knitr::knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage[table,dvipsnames]{xcolor}', x, fixed = TRUE)})
knitr::opts_chunk$set(echo = FALSE,comment = NA, warning = FALSE, message = FALSE, cache = FALSE, fig.height = 3)
setwd("~/Desktop/MoWater/boulder/src")
rm( list = ls() )

library(tidyverse)
library(fields)
library(lubridate)
library(latex2exp)
library(reshape2)
library(gridExtra)
library(viridis)
library(RColorBrewer)

# make sure src is the current working directory

load("../data/final-data.rda")
load("../data/boulderMoWater.rda")
theme_set(theme_bw())
theme_update(panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

\begin{titlepage}
    \newgeometry{top=1in,bottom=1in,left=1in,right=1in,includefoot,includehead}
    \begin{doublespace}
        \centering
        \hrulefill\\[0.25cm]
        {\Huge City of Boulder}\\
        {\Large Exploratory Data Analysis}\\
        \hrulefill
        \vfill
        \includegraphics[width=\linewidth,height=\textheight,keepaspectratio]{"../images/group_photo.png"}\\
        {\large \hfill Justin Nichols \qquad \qquad \qquad \qquad  Arlyn Alcid \qquad \qquad \qquad \qquad Ocean Wave \hfill}
        \vfill
        \includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{"../images/mowater.png"}\hfill
        \includegraphics[width=0.10\linewidth,height=\textheight,keepaspectratio]{"../images/bu.png"}\hfill
        \includegraphics[width=0.15\linewidth,height=\textheight,keepaspectratio]{"../images/mines.png"}\hfill
        \includegraphics[width=0.15\linewidth,height=\textheight,keepaspectratio]{"../images/nsf.png"}
    \end{doublespace}
\end{titlepage}

\newpage

\restoregeometry
\pagestyle{fancy}


We begin by looking at how many observations we have based on the coagulant:

```{r}
  table(final_data$coagulant)
```

The following plots provide visuals for the effluent OP. They are split up by ferric or alum, as well as a comparison of the hourly data after removal with the original 10-minute intervals.

```{r}
  p1 <- alum_data %>% ggplot() + 
    geom_point(aes(date, op_conc_mg_p_l_hourly)) + 
    labs(title = "Alum OP Hourly",
         x = "Date",
         y = "Hourly Ortho-phosphate (mg/L)") + 
    theme(plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 8))
```

```{r}
  alumphos <- phosfax_10m %>% # phosfax only alum coagulant every 10 minutes 
    filter(date >= ymd("2019-10-22")) %>%
    filter(date <= ymd("2019-12-15"))

  rects1 <- data.frame(xstart = as.POSIXct('2019-11-06 15:00:00'), 
                    xend = as.POSIXct('2019-12-01 00:00:00'))
####### thursday plot 1
  p2 <- alumphos %>% ggplot() + 
    geom_point(aes(date, op_conc_mg_p_l), size=1.25) + 
    labs(title = "Alum Effluent OP 10-minute Intervals",
         x = "Date",
         y = "Ortho-phosphate (mg/L)") + 
    geom_rect(data = rects1, aes(xmin = xstart, xmax = xend, 
                                ymin = -Inf, ymax = Inf),fill = "red",col = "red", alpha = 0.1) + 
    theme(plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 8))
  grid.arrange(p1, p2, ncol = 2)
```
\vfill
```{r}
  p1 <- ferr_data %>% ggplot() + 
    geom_point(aes(date, op_conc_mg_p_l_hourly)) + 
    labs(title = "Ferric OP Hourly",
         x = "Date",
         y = "Hourly Ortho-phosphate (mg/L)") + 
    theme(plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 8))
```

```{r}
  ferrphos <- phosfax_10m %>%
    filter(date >= ymd("2019-08-15")) %>%
    filter(date <= ymd("2019-09-07"))
  
  rects2 <- data.frame(xstart = as.POSIXct('2019-08-30 15:00:00'), 
                       xend = as.POSIXct('2019-09-02 10:00:00'))
  
  #### thursday plot 2
  p2 <- ferrphos %>% ggplot() + 
    geom_point(aes(date, op_conc_mg_p_l), size=1.25) + 
    labs(title = "Ferric Effluent OP 10-minute Intervals",
         x = "Date",
         y = "Ortho-phosphate (mg/L)") + 
    geom_rect(data = rects2, aes(xmin = xstart, xmax = xend, 
                                ymin = -Inf, ymax = Inf),fill = "red",col = "red", alpha = 0.1) + 
    theme(plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 8))
  
  grid.arrange(p1, p2, ncol = 2)
```


```{r}
  ferr_op_melt <- melt(ferr_data[,c(1,3:4)], id = "date")
  p1 <- ferr_op_melt %>% ggplot() + 
    geom_point(aes(date, value, col = variable)) +  
    labs(title = "Ferric OP Concentration at Dosing",
         x = "Date",
         y = "Hourly OP (mg/L)") +
    scale_color_brewer(palette = "Dark2", name = "Location", labels = c("Dosing Station", "Effluent")) +
    theme(legend.title = element_text(size = 8),
          legend.text = element_text(size = 7),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = c(0.24,0.2),
          legend.background = element_blank(),
          legend.key = element_blank())
  
  alum_op_melt <- melt(alum_data[,c(1,3:4)], id = "date")
  p2 <- alum_op_melt %>% ggplot() + 
    geom_point(aes(date, value, col = variable)) +  
    labs(title = "Alum OP Concentration at Dosing",
         x = "Date",
         y = "Hourly OP (mg/L)") +
    scale_color_viridis(discrete = TRUE, name = "Location", labels = c("Dosing Station", "Effluent")) +
    theme(legend.title = element_text(size = 8),
          legend.text = element_text(size = 7),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = c(0.67,0.75),
          legend.background = element_blank(),
          legend.key = element_blank())
  
  grid.arrange(p1, p2, ncol = 2)
```
\vfill
```{r}
  p1 <- ferr_data %>% ggplot(aes(op_mg_p_l, op_conc_mg_p_l_hourly)) + 
    geom_point() +
    stat_smooth(method="lm", se=FALSE) + 
    labs(title = "Ferric Effluent vs. Dosing OP Concentrations",
         x = "Dosing OP (mg/L)",
         y = "Effluent OP (mg/L)") +
    theme(plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 8))
  p2 <- alum_data %>% ggplot(aes(op_mg_p_l, op_conc_mg_p_l_hourly)) + 
    geom_point() +
    stat_smooth(method="lm", se=FALSE) + 
    labs(title = "Alum Effluent vs. Dosing OP Concentrations",
         x = "Dosing OP (mg/L)",
         y = "Effluent OP (mg/L)") +
    theme(plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 8))
  
  grid.arrange(p1, p2, ncol = 2)
```
\newpage
```{r}
  ferr_data <- ferr_data %>% mutate(hour = hour(date))
  alum_data <- alum_data %>% mutate(hour = hour(date))
  
  alum_melt1 <- melt(alum_data[,c(8,15)], id.vars = "hour")
  alum_melt2 <- melt(alum_data[,c(13,15)], id.vars = "hour")
  alum_melt3 <- melt(alum_data[,c(9:10,15)], id.vars = "hour")
  alum_melt4 <- melt(alum_data[,c(6,11,15)], id.vars = "hour")
  alum_melt5 <- melt(alum_data[,c(3:4,15)], id.vars = "hour")
  alum_melt6 <- melt(alum_data[,c(5,7,12,14,15)], id.vars = "hour")
  
  ferr_melt1 <- melt(alum_data[,c(8,15)], id.vars = "hour")
  ferr_melt2 <- melt(alum_data[,c(13,15)], id.vars = "hour")
  ferr_melt3 <- melt(alum_data[,c(9:10,15)], id.vars = "hour")
  ferr_melt4 <- melt(alum_data[,c(6,11,15)], id.vars = "hour")
  ferr_melt5 <- melt(alum_data[,c(3:4,15)], id.vars = "hour")
  ferr_melt6 <- melt(alum_data[,c(5,7,12,14,15)], id.vars = "hour")
  hours <- c("7" = "Hour 7", "8" = "Hour 8", "9" = "Hour 9", "10" = "Hour 10", "11" = "Hour 11", "12" = "Hour 12",
             "13" = "Hour 13", "14" = "Hour 14","15" = "Hour 15","16" = "Hour 16","17" = "Hour 17","18" = "Hour 18")
```

```{r, fig.height = 10}
  ggplot(alum_melt1) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(alum_melt2) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(alum_melt3) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(alum_melt4) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(alum_melt5) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(alum_melt6) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(ferr_melt1) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(ferr_melt2) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(ferr_melt3) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(ferr_melt4) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(ferr_melt5) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r, fig.height = 10}
  ggplot(ferr_melt6) +
    geom_boxplot(aes(y=value, fill=variable), outlier.size = 0.5, outlier.color = "red") + 
    facet_wrap(vars(hour), labeller = labeller(hour = hours)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 9),
          plot.title = element_text(face="bold", size = 9, hjust = 0.5),
          axis.text = element_text(size = 6),
          axis.title = element_text(size = 9),
          strip.text.x = element_text(size = 8),
          legend.position = "bottom")
```
\newpage
```{r}
  ferr_diff_pct <- (ferr_data$op_mg_p_l - ferr_data$op_conc_mg_p_l_hourly)/ferr_data$op_mg_p_l * 100
  alum_diff_pct <- (alum_data$op_mg_p_l - alum_data$op_conc_mg_p_l_hourly)/alum_data$op_mg_p_l * 100
```
Percent Change in Alum OP:
```{r}
  alum_diff_pct
```
Percent Change in Ferric OP:
```{r}
  ferr_diff_pct
```
We perform a t-test to determine if they're performing differently
```{r}
  # shows that alum can be concluded as better
  t.test(alum_diff_pct, ferr_diff_pct)
```






